### 2021-05-26 paper/referee_response.md
First adds after referee report

# Referee response

We thank the referee for their precious work and for the comment they made.
We read carefully throught the proposed changes and we amended the text accordingly. The referee can find below thes answers to their questions/concerns

## PnP comments

- Please add citations to the LIGO and Virgo instrument papers the first time these detectors are mentioned (1411.4547, 1408.3978).
		> Done: page 9

- Please fix the broken citation at the beginning of section IV to "the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog”
		> Done, added also the link to DCC entry where the PSD are taken from

- Please link to the correct citation at the start of Section VB, as [27] links to the climate data. "We focus on the public data released by the LIGO/Virgo collaboration [27]. We reconstruct the PSD both assuming the CAT and FPE loss functions on 1000s of data [27] from the Livingston observatory starting from GPS time 1164603392. The data are sampled at 4096 Hz.”
		> Done

- "Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.” Here the Bayeswave papers should be cited (1410.3835, 1410.3852) as well as other key LIGO papers describing glitches, their properties, and subtraction (1508.07316, 1602.03844, 1808.03619), and I believe [28] should point to the Zevin et. al paper in [29].
		> Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. V.B, page 14

- "for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA,” Please again add the citations for the LIGO and Virgo detector papers, and also add a citation to the KAGRA detector papers (1306.6747, 1111.7185).
		> Done: pag 14

- "hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD”: Please add citations to other papers from LVC members discussing marginalization over PSD detector noise (0804.3853, 1109.0442, 1909.01934, 1409.7215, 1307.8195, 2004.05149, 2006.05292, 2101.01200, 2004.07515, 1506.00185 and the two Bayeswave papers mentioned above).
		> The relevant part has been changed so as to acknowledge previous works. Probably here a _misunderstanding_ on what marginalization means?
			Everything is at page 15: "several approaches have been developed to try to study the validity of the assumption or even to drop it [46–52]"

- Please add a citation to https://arxiv.org/abs/gr-qc/0011041, which also demonstrates the applicability of autoregressive processes to gravitational-wave power spectral density estimation.
		> Done: on sec. V.B, page 13

- I’m not sure if this qualifies as a PnP comment, but a significant omission is the lack of discussion of the Bayeswave method of PSD estimation, which is the official method for calculating PSDs for parameter estimation results by the LVC. In addition to citing the papers described above, I think it would be appropriate to add a discussion, either in the Introduction or the Final Remarks comparing MESA to this “on-source” PSD estimation and citing the relvant papers disucssing the Bayeswave PSD method in comparison to the Welch method, some repeated from the above requests (1907.06540, 2006.05292).
		> This is a very good suggestion... Are we gonna include BayesLine among our comparison? It looks like a good idea. If we do so, probably we want to say something about it in the beginning
			The paper 1907.06540 is already cited together with BayesWave

## Personal comments

We fixed all the typos that the referee pointed out. We are thankful for that.

### Comparison with Welch method:
- Maybe it’s because of the broken citation, but I don’t understand which PSD is being used in the simulation. Is it the same “analytical PSD computed for the LIGO Hanford interferometer” used in the previous section, or is it a PSD calculated using the real LIGO data from around GW150914? If so, how was that PSD calculated?
		> The two PSDs are different. In the validation part, we used the design sensitivity of aLIGO (we quoted the relevant paper as well as the DCC link for the file). In the Welch comparison part, we employed the PSD released together with GW150914 (again we quoted the GWTC-1 paper and the link).

### Applications:

- It’s not really clear to me when you perform the forecasting for the temperature data why CAT is the obvious choice for the loss function, since in the previous section you described how CAT is unable to recover the true order of the autoregressive process. Does that mean that it would be worse at forecasting based on the AR(p)? There seems to be a contradiction where CAT does a better job for the temperature data but worse for the simulation presented in section II. It would be helpful to elaborate on this.
		> We added a claryfing footnote on this (note 10, page 13). We hope it is clear.

- "Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.”: How does this compare to the method presented in https://arxiv.org/abs/1908.05644, where the data containing a glitch are “inpainted” with values chosen such that the inverse-PSD-filtered data is 0 at these times?
		> Work in progress: we should understand what's in the paper...

### Final Remarks and Future Prospects

- In addition to adding the citations to other LVC works presenting methods and results for marginalizing over the PSD uncertainty (see PnP comments above), a more detailed discussion of how MESA compares to these other methods (particularly the error in the PSD measurement) would be appropriate here.
		> Working on it: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?